A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-11 main.tex
Wrote till Typing rules for a multi-stage calculus.

Terms include a constant $c$ which is declared in signature $\Sigma$.
In addition to ordinary terms in Simply Typed Lambda Calculus, there are five more forms.
$\TB_\alpha M$ represents a code fragment, and $\TB_\alpha M$ represents escape.
$\Lambda\alpha.M$ is a stage variable abstraction.
$M\ \epsilon$ is an application of $\epsilon$ to a stage variable.
$\%_\alpha M$ is a primitive operator for cross-stage persistence.
$\TB_\alpha, \TBL_\alpha, \%_\alpha$ tie stronger than the two forms of applications
and applications are left-associative
and two abstractions extends as far to the right as possible.
For example, $\F\alpha.\lambda x:\I.\TB_\alpha x\ y$ means $\F\alpha.(\lambda x:\I.(\TB_\alpha x)\ y)$.

@@

In this section, we define full reduction for \LMD.
Before giving the definition of reduction, we define two substitutions.
Substitution $M[x\mapsto N]$ is the normal capture-avoiding substitution, and we omit its definition here.
Substitution $M[\alpha \mapsto \epsilon]$ is defined below.

@@

  In this section, we define the type system of \LMD.
  \LMD\ is little complicated because it contains dependent types.
  It contains typing, kinding, well-formed kinding, term equality, type equality, and kind equality.
  In other words, there are six types of judgement in \LMD.
  We show them in Figure \ref{fig:LMD-six-judgements}.

  \begin{figure}
      \begin{center}
          \begin{align*}
              \G & \V M : \tau @ A   \\
              \G & \V \tau :: K @ A  \\
              \G & \V K \iskind @ A  \\
              \G & \V M \E N @ A     \\
              \G & \V \tau \E \sigma \\
              \G & \V K \E J @ A     
          \end{align*}
          \caption{Six types of judgement in \LMD}
          \label{fig:LMD-six-judgements}
      \end{center}
  \end{figure}

  Especially, we discuss on the type equality of \LMD in detail
  because it is the main contribution of this paper.
  \LMD is developed on the basis of \LTP but there are no type equivalence rules in \LTP.

  \begin{definition}[Typing]
      The typing relation $ \G \V M : \tau @ A $ is the least relation closed under the rules in Figure \ref{fig:typing-rules}.
  \end{definition}

  % 各論

  % 通常のルール
  We show typing rules of \LMD in Figure \ref{fig:typing-rules}.
  The rules \TVar , \TAbs, and \TApp\ are almost the same as those in simply typed lambda calculus 
  if you ignore stage annotations and kind checking.
  The rule \TVar\ means that a variable can appear only at the stage in which it is declared.
  It checks the type of variable $x$, $\tau$, has the proper kind for terms, $*$.
  \TAbs\ also checks the kind of the type.

  The rule \TConst\ means any constants in the signature can appear at any stage.
  For example, if we have a signature $\Sigma$ which is $\textrm{bool} :: *, \textrm{true}: \textrm{bool}, \textrm{false}: \textrm{bool}$,
  the derivation tree in Figure \ref{fig:tconst-derivation-tree} is admissible.

@@

  % 多段階計算のルール
  The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp\ are rules for a multi-stage calculus.
  They are corresponding to quoting a code, unquoting a code, making a stage abstraction, 
  application of $\epsilon$ stage to a stage abstraction, and cross-stage persistence, respectively.
  Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

@@

  Strong Normalization is also important property which guarantee that
  no typed term has an infinite reduction sequence.
  We prove this thorem by translating \LMD to Simply Typed Lambda Calculus.

@@

  \begin{proof}
      In order prove this theorem, we define a translation $\natural$ from \LMD\ to Simply Typed Lambda Calculus.
      Second, we prove $\natural$ preserves typing and $\beta$-reductions.
      Then, we can prove Strong Normalization of \LMD\ from Strong Normalization of Simply Typed Lambda Calculus.
  \end{proof}

@@

  % \section{First Section}
  % \subsection{A Subsection Sample}
  % Please note that the first paragraph of a section or subsection is
  % not indented. The first paragraph that follows a table, figure,
  % equation etc. does not need an indent, either.
  % 
  % Subsequent paragraphs, however, are indented.
  % 
  % \subsubsection{Sample Heading (Third Level)} Only two levels of
  % headings should be numbered. Lower level headings remain unnumbered;
  % they are formatted as run-in headings.
  % 
  % \paragraph{Sample Heading (Fourth Level)}
  % The contribution should contain no more than four levels of
  % headings. Table~\ref{tab1} gives a summary of all heading levels.
  % 
  % \begin{table}
  % \caption{Table captions should be placed above the
  % tables.}\label{tab1}
  % \begin{tabular}{|l|l|l|}
  % \hline
  % Heading level &  Example & Font size and style\\
  % \hline
  % Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
  % 1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
  % 2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
  % 3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
  % 4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
  % \hline
  % \end{tabular}
  % \end{table}
  % 
  % 
  % \noindent Displayed equations are centered and set on a separate
  % line.
  % \begin{equation}
  % x + y = z
  % \end{equation}
  % Please try to avoid rasterized images for line-art diagrams and
  % schemas. Whenever possible, use vector graphics instead (see
  % Fig.~\ref{fig1}).
  % 
  % \begin{figure}
  % \includegraphics[width=\textwidth]{fig1.eps}
  % \caption{A figure caption is always placed below the illustration.
  % Please note that short captions are centered, while long ones are
  % justified by the macro package automatically.} \label{fig1}
  % \end{figure}
  % 
  % \begin{theorem}
  % This is a sample theorem. The run-in heading is set in bold, while
  % the following text appears in italics. Definitions, lemmas,
  % propositions, and corollaries are styled the same way.
  % \end{theorem}
  % %
  % % the environments 'definition', 'lemma', 'proposition', 'corollary',
  % % 'remark', and 'example' are defined in the LLNCS documentclass as well.
  % %
  % \begin{proof}
  % Proofs, examples, and remarks have the initial word in italics,
  % while the following text appears in normal font.
  % \end{proof}
  % For citations of references, we prefer the use of square brackets
  % and consecutive numbers. Citations using labels or the author/year
  % convention are also acceptable. The following bibliography provides
  % a sample reference list with entries for journal
  % articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
  % book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
  % and a homepage~\cite{ref_url1}. Multiple citations are grouped
  % \cite{ref_article1,ref_lncs1,ref_book1},
  % \cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
  %
  % ---- Bibliography ----
  %
  % BibTeX users should specify bibliography style 'splncs04'.
  % References will then be sorted and formatted in the correct style.
  %
  \bibliographystyle{splncs04}
  \bibliography{main}
  %
  % \begin{thebibliography}{8}
  % \bibitem{ref_article1}
  % Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
  % 
  % \bibitem{ref_lncs1}
  % Author, F., Author, S.: Title of a proceedings paper. In: Editor,
  % F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
  % Springer, Heidelberg (2016). \doi{10.10007/1234567890}
  % 
  % \bibitem{ref_book1}
  % Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
  % Location (1999)
  % 
  % \bibitem{ref_proc1}
  % Author, A.-B.: Contribution title. In: 9th International Proceedings
  % on Proceedings, pp. 1--2. Publisher, Location (2010)
  % 
  % \bibitem{ref_url1}
  % LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
  % Oct 2017
  % \end{thebibliography}

>>>>

Terms include a constant $c$ which is declared in signature $\Sigma$.
In addition to ordinary terms in the symply typed lambda calculus, there are five more forms.
$\TB_\alpha M$ represents a code fragment, and $\TB_\alpha M$ represents escape.
$\Lambda\alpha.M$ is a stage variable abstraction.
$M\ \epsilon$ is an application of $\epsilon$ to a stage variable.
$\%_\alpha M$ is a primitive operator for cross-stage persistence.
$\TB_\alpha, \TBL_\alpha, \%_\alpha$ tie stronger than the two forms of applications
and applications are left-associative
and two abstractions extends as far to the right as possible.
For example, $\F\alpha.\lambda x:\I.\TB_\alpha x\ y$ means $\F\alpha.(\lambda x:\I.(\TB_\alpha x)\ y)$.

@@

Next, we define full reduction for \LMD.
Before giving the definition of reduction, we define two substitutions.
Substitution $M[x\mapsto N]$ is the normal capture-avoiding substitution, and we omit its definition here.
Substitution $M[\alpha \mapsto \epsilon]$ is defined below.

@@

In this section, we define the type system of \LMD.
\LMD is little complicated because it contains dependent types.
It contains typing, kinding, well-formed kinding, term equality, type equality, and kind equality.
In other words, there are six types of judgement in \LMD.
We show them in Figure \ref{fig:LMD-six-judgements}.

\begin{figure}
	\begin{center}
		\begin{align*}
			\G & \V M : \tau @ A   & \text{typing} \\
			\G & \V \tau :: K @ A  & \text{kinding} \\
			\G & \V K \iskind @ A  & \text{well-formed kinding} \\
			\G & \V M \E N @ A     & \text{term equality} \\
			\G & \V \tau \E \sigma & \text{type equality} \\
			\G & \V K \E J @ A     & \text{kind equality}
		\end{align*}
		\caption{Six types of judgement in \LMD}
		\label{fig:LMD-six-judgements}
	\end{center}
\end{figure}

\begin{definition}[Typing]
	The typing relation $ \G \V M : \tau @ A $ is the least relation closed under the rules in Figure \ref{fig:typing-rules}.
\end{definition}
We start from defining typing of \LMD.
Typing rules of \LMD are shown in Figure \ref{fig:typing-rules}.
The rule \TConst\ means any constants in the signature can appear at any stage.
For example, if we have a signature $\Sigma$ which is $\textrm{bool} :: *, \textrm{true}: \textrm{bool}, \textrm{false}: \textrm{bool}$,
the derivation tree in Figure \ref{fig:tconst-derivation-tree} is admissible.
The rules \TVar , \TAbs, and \TApp\ are almost the same as those in the simply typed lambda calculus 
if you ignore stage annotations and kind checking.
The rule \TVar\ means that a variable can appear only at the stage in which it is declared.
It checks the type of variable $x$, $\tau$, has the proper kind for terms, $*$.
The rule \TAbs\ also checks the kind of the type.
The rule \TApp\ requires $M$ and $N$ are typed at the same stage.

@@

% Typing rules for a multi-stage calculus
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp\ are rules for a multi-stage calculus.
They are corresponding to brackets, escaping a code, making a stage abstraction, 
application of $\epsilon$ stage to a stage abstraction, and cross-stage persistence, respectively.
Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

@@

Strong Normalization is also important property which guarantee that
no typed term has an infinite reduction sequence.
We prove this thorem by translating \LMD to the symply typed lambda calculus.

@@

\begin{proof}
	In order prove this theorem, we define a translation $\natural$ from \LMD\ to the symply typed lambda calculus.
	Second, we prove $\natural$ preserves typing and $\beta$-reductions.
	Then, we can prove Strong Normalization of \LMD\ from Strong Normalization of the symply typed lambda calculus.
\end{proof}

@@

% \section{First Section}
% \subsection{A Subsection Sample}
% Please note that the first paragraph of a section or subsection is
% not indented. The first paragraph that follows a table, figure,
% equation etc. does not need an indent, either.
% 
% Subsequent paragraphs, however, are indented.
% 
% \subsubsection{Sample Heading (Third Level)} Only two levels of
% headings should be numbered. Lower level headings remain unnumbered;
% they are formatted as run-in headings.
% 
% \paragraph{Sample Heading (Fourth Level)}
% The contribution should contain no more than four levels of
% headings. Table~\ref{tab1} gives a summary of all heading levels.
% 
% \begin{table}
% \caption{Table captions should be placed above the
% tables.}\label{tab1}
% \begin{tabular}{|l|l|l|}
% \hline
% Heading level &  Example & Font size and style\\
% \hline
% Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
% 1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
% 2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
% 3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
% 4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
% \hline
% \end{tabular}
% \end{table}
% 
% 
% \noindent Displayed equations are centered and set on a separate
% line.
% \begin{equation}
% x + y = z
% \end{equation}
% Please try to avoid rasterized images for line-art diagrams and
% schemas. Whenever possible, use vector graphics instead (see
% Fig.~\ref{fig1}).
% 
% \begin{figure}
% \includegraphics[width=\textwidth]{fig1.eps}
% \caption{A figure caption is always placed below the illustration.
% Please note that short captions are centered, while long ones are
% justified by the macro package automatically.} \label{fig1}
% \end{figure}
% 
% \begin{theorem}
% This is a sample theorem. The run-in heading is set in bold, while
% the following text appears in italics. Definitions, lemmas,
% propositions, and corollaries are styled the same way.
% \end{theorem}
% %
% % the environments 'definition', 'lemma', 'proposition', 'corollary',
% % 'remark', and 'example' are defined in the LLNCS documentclass as well.
% %
% \begin{proof}
% Proofs, examples, and remarks have the initial word in italics,
% while the following text appears in normal font.
% \end{proof}
% For citations of references, we prefer the use of square brackets
% and consecutive numbers. Citations using labels or the author/year
% convention are also acceptable. The following bibliography provides
% a sample reference list with entries for journal
% articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
% book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
% and a homepage~\cite{ref_url1}. Multiple citations are grouped
% \cite{ref_article1,ref_lncs1,ref_book1},
% \cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{main}
%
% \begin{thebibliography}{8}
% \bibitem{ref_article1}
% Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
% 
% \bibitem{ref_lncs1}
% Author, F., Author, S.: Title of a proceedings paper. In: Editor,
% F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
% Springer, Heidelberg (2016). \doi{10.10007/1234567890}
% 
% \bibitem{ref_book1}
% Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
% Location (1999)
% 
% \bibitem{ref_proc1}
% Author, A.-B.: Contribution title. In: 9th International Proceedings
% on Proceedings, pp. 1--2. Publisher, Location (2010)
% 
% \bibitem{ref_url1}
% LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
% Oct 2017
% \end{thebibliography}
